---
title: "dive-in-minor-functions-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dive-in-minor-functions-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(wastewateR)

```


This vignette displays a step-by-step example of how the minor functions act and what messaging the user can expect from each individual piece.


## Step 1: Look at Data, Format Correctly

The wastewateR package processing functions expect the data frame supplied to them to have the following columns: 

<b style='color:#DB5461;'>
- “Well”, “Sample”, “Target”, “Concentration”, “CopiesPer20uLWell”, “AcceptedDroplets”, “Positives”, and “Negatives”
</b>

<em>
Depending on your output, the columns in your data may also be: 

- "Well", "Sample.description.1", "Target", "Conc.copies.µL.", "Copies.20µLWell", "Accepted.Droplets", "Positives", and "Negatives"

In that case, you will need to rename them to align with the expected naming convention.
</em>

```{r}

n1_example_data <- read.csv("n1_example.csv") 

# what columns are in my data set? 
print(colnames(n1_example_data))

# what Target does this data apply to?
print(unique(n1_example_data$Target))

```

From the above, we see that we have too many columns, and that we have SARS-CoV-2 N1 data. Let's first select only the columns we need for QAQC processing of this data.

```{r}

n1_example_data <- n1_example_data %>% select(Well, Sample, Target, Concentration, CopiesPer20uLWell, AcceptedDroplets, Positives, Negatives)

```

Then, let's take a look at our "plate" to understand how this data looks even further. 

::: {.panel-tabset}

### Concentration

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = Concentration)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "Concentration") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

### CopiesPer20uLWell

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = CopiesPer20uLWell)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "CopiesPer\n20uLWell") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

### AcceptedDroplets

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = AcceptedDroplets)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "Accepted\nDroplets") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

### Positives

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = Positives)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "Positives") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

### Negatives

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = Negatives)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "Negatives") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

:::


We can see that this is not a full plate of data, as in a typical 96 well plate, there are rows of A-I and columns of 1-12. That's a-okay, and will not affect the functioning of this R package. 

We can also see that we have 15 Samples in triplicate, a positive control (POS) set in triplicate, a negative control (NEG) set in triplicate, a non-template control (NTC) set in triplicate, and an extraction control (EXT) set in triplicate, for a total of 57 wells of data.

While we could just apply the wrapper function `qaqc_processing_sc2_n1_slim()` to this data, below we will walk through each minor function that makes up that wrapper function while making deliberate edits to our data file. This will allow us to see what the messaging and output of each minor function looks like. If the user will only be using the wrapper functions, the messaging and final output will be the two things that are most relevant. If the user will be building their own set of functions or wrapper function to process their data, this guide will hopefully address common questions, and all portions would be relevant.  

## Step 2: Individual Minor Functions

#### w0110_sample_name_edits()

#### w0115_target_check()

#### w0125_expected_controls_present()

#### w0150_sample_naming_structure()

#### w0200_accepted_droplet_count()

#### w0300_ntc_control_check()

#### w0400_pos_control_hard_stop()

#### w0450_pos_control_breakdown_check()

#### w0600_ext_neg_control_check()

#### w0650_cumulative_count_check()

#### w0700_pos_droplet_sum()

#### w0750_neg_droplet_samples()

#### w0800_recover_control_check()

#### w0900_positives_comparison_rule()

#### w1000_remove_rows_as_chosen()





