---
title: "dive-in-minor-functions-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{dive-in-minor-functions-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(wastewateR)

```


This vignette displays a step-by-step example of how the minor functions act and what messaging the user can expect from each individual piece.


## Step 1: Look at Data, Format Correctly

The wastewateR package processing functions expect the data frame supplied to them to have the following columns: 

<b style='color:#DB5461;'>
- “Well”, “Sample”, “Target”, “Concentration”, “CopiesPer20uLWell”, “AcceptedDroplets”, “Positives”, and “Negatives”
</b>

<em>
Depending on your output, the columns in your data may also be: 

- "Well", "Sample.description.1", "Target", "Conc.copies.µL.", "Copies.20µLWell", "Accepted.Droplets", "Positives", and "Negatives"

In that case, you will need to rename them to align with the expected naming convention.
</em>

```{r}

n1_example_data <- read.csv("n1_example.csv") 

# what columns are in my data set? 
print(colnames(n1_example_data))

# what Target does this data apply to?
print(unique(n1_example_data$Target))

```

From the above, we see that we have too many columns, and that we have SARS-CoV-2 N1 data. Let's first select only the columns we need for QAQC processing of this data.

```{r}

n1_example_data <- n1_example_data %>% select(Well, Sample, Target, Concentration, CopiesPer20uLWell, AcceptedDroplets, Positives, Negatives)

```

Then, let's take a look at our "plate" to understand how this data looks even further. 

::: {.panel-tabset}

### Concentration

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = Concentration)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "Concentration") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

### CopiesPer20uLWell

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = CopiesPer20uLWell)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "CopiesPer\n20uLWell") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

### AcceptedDroplets

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = AcceptedDroplets)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "Accepted\nDroplets") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

### Positives

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = Positives)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "Positives") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

### Negatives

```{r}

ggplot() + 
  geom_tile(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), fill = Negatives)) + 
  geom_text(data = n1_example_data, aes(x = substr(Well, 2, 3), y = substr(Well, 1, 1), label = paste0("Sample:\n", Sample, "\nTarget:\n", Target)), size = 3) + 
  scale_y_discrete(limits=rev) + 
  scale_x_discrete(position = "top") + 
  labs(x = "", 
       y = "", 
       fill = "Negatives") + 
  theme_bw() + 
  scale_fill_gradient(low = "#C7E6DD", high = "#EA9AA2")

```

:::



